Generative AI for image reconstruction in Intensity Interferometry: a first attempt

### 2025-07-16 Phase_II/draft/prev/GAN.tex
edits

\section{Generative Adversarial Networks}
Generative Adversarial Networks (GANs) were introduced by \cite{goodfellow2014generative}. The underlying concept is straightforward: it involves two competing networks. The first network, known as the Generator, produces new images based on an input image. These will be referred to as generated images. The second network, the Discriminator, attempts to distinguish between the generated image (predicted image) and the real image (ground truth).

Through the alternating training of these networks, the generated images gradually become indistinguishable from the real images. Essentially, this process constitutes a two-player min-max game --- a classic problem in game theory. The original formulation of GANs is given by:
\begin{equation}
	\centering
	\begin{aligned}
		\min_{G} \max_{D} V(D, G) &= \mathbb{E}_{x \sim p_{\rm data}(x)} \left[ \log D(x) \right] \\
		&+ \mathbb{E}_{z \sim p_{z}(z)} \left[ \log \left( 1-D(G(z)) \right) \right]
	\end{aligned}
	\label{eq:Basic_GAN}
\end{equation}
where $V(D, G)$ denotes the value function of the min-max game.

@@

		\max_{D} V(G, D) &= \mathbb{E}_{x \sim p_{\rm data}} \left[ \log D^{*}_{G}(x) \right] \\ 

@@

where \(D^{*}_{G}\) denotes the optimum of the discriminator for a given fixed generator, as shown in equation (\ref{eq:Disc_optimum}). It can be demonstrated that the global optimum of equation (\ref{eq:GAN_reformulated}) is achieved if and only if \(p_G = p_{\rm data}\). Furthermore, if both \(G\) and \(D\) are allowed to reach their respective optima, then \(p_G\) converges to \(p_{\rm data}\). A more comprehensive discussion of the problem, including proofs, is provided in \cite{goodfellow2014generative}.

@@

	D^*_G(x) = \frac{p_{\rm data}(x)}{p_{\rm data}(x) + p_g(g)}

@@

Subsequently, the GAN framework was extended to a conditional model \citep{mirza2014conditional}. In this formulation, both the Generator and the Discriminator receive additional information $y$, and the value function of the conditional GAN (cGAN) is expressed as:
\begin{equation}
	\centering
	\begin{aligned}
		V(D, G) &= \mathbb{E}_{x \sim p_{\rm data}(x)} \left[ \log D(x|y) \right] \\
		&+ \mathbb{E}_{z \sim p_{z}(z)} \left[ \log \left( 1-D(G(z|y) \right) \right]
	\end{aligned}
	\label{eq:conditional_GAN}
\end{equation}
\cite{isola2017image} further observed that combining the cGAN from Eq.~\eqref{eq:conditional_GAN} with the traditional L1 loss improves the results, as the Generator is encouraged to produce outputs closer to the ground truth. Hence, the function that is minimized is:
\begin{equation}
	\centering
	L_{tot} = \arg \min_{G} \max_{D} V(D, G) + \lambda \cdot L_1(G)
	\label{eq:total_gen_loss}
\end{equation}
with the choice $\lambda = 100$ and
\begin{equation}
	\centering
	L_1(G) = \mathbb{E}_{x,y,z} \left[ ||{y - G(x,z)}||_1 \right]
	\label{eq:l1_loss}
\end{equation}
This type of network has demonstrated remarkable robustness across a variety of applications. For example, it can generate colored images from grayscale inputs based on architectural labels, transform images from day to night, and even predict maps from satellite data. A more extensive list of applications is provided in \cite{isola2017image}.

@@

In contrast, the upsampling process uses only the standard Rectified Linear Unit (ReLU) for neuron activation. This stage also comprises convolutional layers followed by operations with a stride of 2 to upscale the image to a higher resolution. Additionally, a dropout layer is introduced at the beginning of the upsampling phase to mitigate overfitting of the Generator model \citep{isola2017image}.

@@

\subsection{Discriminator}
As discussed earlier, the Discriminator is tasked with classifying the images produced by the Generator as either real or fake. It takes a real image from the dataset (often referred to as the target image for the Generator) and provides feedback to guide the Generator toward producing more accurate images. In this work, the PatchGAN model \citep{isola2017image} is employed as the Discriminator. Unlike a traditional global classifier, PatchGAN evaluates individual patches of the image, outputting a grid of predictions rather than a single scalar value.

>>>>

\section{Generative Adversarial Networks}
Generative Adversarial Networks (GANs) were introduced by \citep{goodfellow2014generative}. The underlying concept is straightforward: it involves two competing networks. The first network, known as the Generator, produces new images based on an input image. These will be referred to as generated images. The second network, the Discriminator, attempts to distinguish between the generated image (predicted image) and the real image (ground truth).

Through the alternating training of these networks, the generated images gradually become indistinguishable from the real images. Essentially, this process constitutes a two-player min-max game --- a classic problem in game theory. The original formulation of GANs is given by:
\begin{equation}
	\centering
	\begin{aligned}
		\min_{G} \max_{D} V(D, G) &= \mathbb{E}_{x \sim p_{data}(x)} \left[ \log D(x) \right] \\
		&+ \mathbb{E}_{z \sim p_{z}(z)} \left[ \log \left( 1-D(G(z)) \right) \right]
	\end{aligned}
	\label{eq:Basic_GAN}
\end{equation}
where $V(D, G)$ denotes the value function of the min-max game.

@@

		\max_{D} V(G, D) &= \mathbb{E}_{x \sim p_{data}} \left[ \log D^{*}_{G}(x) \right] \\ 

@@

where \(D^{*}_{G}\) denotes the optimum of the discriminator for a given fixed generator, as shown in equation (\ref{eq:Disc_optimum}). It can be demonstrated that the global optimum of equation (\ref{eq:GAN_reformulated}) is achieved if and only if \(p_G = p_{data}\). Furthermore, if both \(G\) and \(D\) are allowed to reach their respective optima, then \(p_G\) converges to \(p_{data}\). A more comprehensive discussion of the problem, including proofs, is provided in \cite{goodfellow2014generative}.

@@

	D^*_G(x) = \frac{p_{data}(x)}{p_{data}(x) + p_g(g)}

@@

Subsequently, the GAN framework was extended to a conditional model \citep{mirza2014conditional}. In this formulation, both the Generator and the Discriminator receive additional information $y$, and the value function of the conditional GAN (cGAN) is expressed as:
\begin{equation}
	\centering
	\begin{aligned}
		V(D, G) &= \mathbb{E}_{x \sim p_{data}(x)} \left[ \log D(x|y) \right] \\
		&+ \mathbb{E}_{z \sim p_{z}(z)} \left[ \log \left( 1-D(G(z|y) \right) \right]
	\end{aligned}
	\label{eq:conditional_GAN}
\end{equation}

\cite{isola2017image} further observed that combining the cGAN from equation \ref{eq:conditional_GAN} with the traditional L1 loss improves the results, as the Generator is encouraged to produce outputs closer to the ground truth. Hence, the function that is minimized is:
\begin{equation}
	\centering
	L_{tot} = \arg \min_{G} \max_{D} V(D, G) + \lambda \cdot L_1(G)
	\label{eq:total_gen_loss}
\end{equation}
where $\lambda$ = 100 and $L_1(G)$ is given as
\begin{equation}
	\centering
	L_1(G) = \mathbb{E}_{x,y,z} \left[ ||{y - G(x,z)}||_1 \right]
	\label{eq:l1_loss}
\end{equation}

This type of network has demonstrated remarkable robustness across a variety of applications. For example, it can generate colored images from grayscale inputs based on architectural labels, transform images from day to night, and even predict maps from satellite data. A more extensive list of applications is provided in \cite{isola2017image}.

@@

In contrast, the upsampling process uses only the standard Rectified Linear Unit (ReLU) for neuron activation. This stage also comprises convolutional layers followed by operations with a stride of 2 to upscale the image to a higher resolution. Additionally, a dropout layer is introduced at the beginning of the upsampling phase to mitigate overfitting of the Generator model \cite{isola2017image}.

@@

\subsection{Discriminator}
As discussed earlier, the Discriminator is tasked with classifying the images produced by the Generator as either real or fake. It takes a real image from the dataset (often referred to as the target image for the Generator) and provides feedback to guide the Generator toward producing more accurate images. In this work, the PatchGAN model \cite{isola2017image} is employed as the Discriminator. Unlike a traditional global classifier, PatchGAN evaluates individual patches of the image, outputting a grid of predictions rather than a single scalar value.

### 2025-07-16 Phase_II/draft/prev/main.tex
edits

\author[Rai et al.]{Km Nitu Rai,$^{1,2}$ Yuri van der Burg,$^{3}$ Soumen Basak,$^{1}$ Prasenjit Saha,$^{3}$ and
	Subrata Sarangi$^{4}$ \\ \\
	$^{1}$School of Physics, Indian Institute of Science Education and Research Thiruvananthapuram, Maruthamala PO,
	Vithura,\\ Thiruvananthapuram 695551, Kerala, India \\
	$^{2}$Aryabhatta Research Institute of Observational Sciences, Manora Peak, Nainital 263129, India. \\
	$^{3}$Physik-Institut, University of Zurich, Winterthurerstrasse 190, 8057 Zurich, Switzerland \\
	$^{4}$School of Applied Sciences, Centurion University of Technology and Management, Odisha-752050, India.}

>>>>

\author[Rai et al.]{Km Nitu Rai,$^{1,2}$ Yuri van der Burg,$^{3}$ Soumen Basak,$^{1}$ Prasenjit Saha,$^{3}$ and
	Subrata Sarangi$^{4,5}$ \\ \\
	$^{1}$School of Physics, Indian Institute of Science Education and Research Thiruvananthapuram, Maruthamala PO,
	Vithura,\\ Thiruvananthapuram 695551, Kerala, India \\
	$^{2}$Aryabhatta Research Institute of Observational Sciences, Manora Peak, Nainital 263129, India. \\
	$^{3}$Physik-Institut, University of Zurich, Winterthurerstrasse 190, 8057 Zurich, Switzerland \\
	$^{4}$School of Applied Sciences, Centurion University of Technology and Management, Odisha-752050, India. \\
    $^5$ Visiting Associate, Inter-University Centre for Astronomy and Astrophysics, Post Bag 4, Ganeshkhind, Pune 411 007,\\ Maharashtra, India.}
